### 2021-11-30 chapter5/chapter.tex
sec 5.1

Now consider approximating the initial density \(f(x)=1\). Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error, and \(L^2\) error between the true initial distribution and the reconstructions, we now see that the normalised reconstructions outperform the unnormalised reconstruction. This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(x\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
\end{figure}

@@

@@

\section{Travelling wave}
Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with varying initial conditions. Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\). The PDE (if/when it exists) which describes this system is 
\[\cfrac{\partial}{\partial t} f(x,t) = \cfrac{\partial}{\partial x} f(x,t),\]
where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
\[f(x,t) = f(x-t,0)\]
so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).

We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space. We discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We then

>>>>

Now consider approximating the initial density \(f(x)=1\). Observing Figure~\ref{fig: fun 4 ks error qbdrap closing vecs} of the KS error, and \(L^2\) error between the true initial distribution and the reconstructions, we now see that the normalised reconstructions outperform the unnormalised reconstruction. This suggests that, in this case, the `folding' of closing operator about \(\Delta\) has greatly increased the ability of the reconstruction to approximate this initial distribution. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/qbdrap_closing_vec/fun4/l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(x\), and the approximation (left) and \(L^2\) error between the true PDF and the approximation (right) for the three closing vectors considered; unnormalised (blue solid line with crosses), naive normalised (orange dashed line with diamonds) and normalised (green dash-dotted line with circles). The naive normalised (orange) an normalised closing vectors are coincident.}
	\label{fig: fun 4 ks error qbdrap closing vecs}
\end{figure}

@@

Given the evidence above we choose to use the normalised closing operator to reconstruct solutions. Further, unlike the naive normalised operator, the normalised operator is theoretically justified in that we proved the closing operator leads to a convergent scheme and is linear.

@@

First we consider the initial condition with CDF \(1(x\geq 0.5)\), that is, a point mass at 0.5. This distribution does not have a PDF so we can compare the CDFs only. In Figure~\ref{fig: fun 1 comp} we plot the KS metric (left) and \(L^1\) metric (right) between the true CDF and the reconstructed approximations. Observing the KS metric, it appears that none of the methods converge and the KS error sits around 0.5. This reflects the fact that convergence in distribution implies pointwise convergence of the CDFs except at points of discontinuity. None of the methods appear to converge at this point. Observing the \(L^1\) error between the true CDF and reconstructed approximation (which is the area between the two CDFs) we now see that the methds appear to show the convergent behaviour we expect. Here, the uniformisation method appears to perform the best, while the QBD-RAP method performs the worst. The rate of convergence of the QBD-RAP method appears to be similar to the rate of convergence of the DG scheme.

Perhaps it is no surprise that the uniformisation method performs best. In the uniformisation method as we increase the order, we partition the cell \([0,1]\) into smaller sub-cells, and use constant functions on each sub-cell to approximate the function. As such, the uniformisation method can produce a piecewise continous, linear approximation to the CDF. In contrast, both the DG and QBD-RAP methods result in a smooth approximation to the CDF. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun1/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun1/meshs_l1_cdf_error_formatted.pdf}
	\caption{KS error between the true CDF, \(1(x\geq 0.5)\), and the approximations (left) and \(L^1\) error between the true CDF and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 1 comp} 
\end{figure}

In Figure~\ref{fig: pdf comp fun 1} we plot the approximated CDFs for the DG, uniformisation and QBD-RAP methods alonside the true CDF. The DG method displays undesireable fetures for an approximation to a CDF -- it is not monotonically increasing, at some points it is negative and at some points it is above 1. On the other hand, although the QBD-RAP method converges slowest, it displays good properties in that it results in a monotonically increasing CDF, starting at 0 and ending at 1. 
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{chapter5/figs/comp/fun1/cdfs_formatted.pdf}
	\caption{Reconstructed CDFs using the DG (blue solid line), uniformisation (orange dashed line) and QBD-RAP (green dot-dashed line) methods. The true distribution function is \(1(x\geq 0.5)\) (grey dotted line).}
	\label{fig: pdf comp fun 1}
\end{figure}

Now consider approximaing the initial distribution with PDF \(1(x\leq 0.5)\). Figure~\ref{fig: fun 2 comp} plots the KS error (left) and \(L^2\) error between the true and approximate PDFs (right). Figure~\ref{fig: fun 2 comp} suggests that all methods converge at a similar rate for this problem. Once again, the QBD-RAP method performs worst, the uniformisation method second, and the DG method the best. However, once again the DG method exhibits undesireable properties -- the approximation to the CDF is, at some points above 1, is not monotonic, though these violations are not as severe in this case as they were for the case above. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun2/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun2/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(1(x\geq 0.5)\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 2 comp} 
\end{figure}

So far we have considered two problems which exhibit discontinuities. At the other extreme we now consider an initial distribution with density \(-6z^2+6x\). In Figure~\ref{fig: fun 6 comp} we plot the KS error and the \(L^2\) error between the true and approximated PDF. Since the DG method projects the initial condition onto a polynomial basis, then, for an order 3 approximation and above, the DG method can theoretically approximate the initial condition exactly. This is reflected in Figure~\ref{fig: fun 6 comp}, where the blue curve drops sharply from 1-3, then plateaus. Due to numerical integration errors, for example in the evaluation of the integral in the \(L^2\) norm, and due to machine arithmetic, the errors for the DG scheme are not 0. Regarding the other two methods, they too appear to be convergent at approximately the same rate, with the uniformisation method perform better for the KS error, but very similarly in terms of the \(L^2\) error. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun6/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun6/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(-6z^2+6x\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 6 comp} 
\end{figure}

Consider now the initial distribution with PDF \(cos(4\pi(x+0.5))+1\). Figure~\ref{fig: fun 9 comp} shows the errors. Both the KS error (left) and \(L^2\) norm between the true and approximated PDFs tell a similar story. For sufficiently high order, the DG method approximates the inital condition very well. The uniformisation method performs second best, while the QBD-RAP method is slowest to converge. 
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun9/meshs_ks_error_formatted.pdf}%
	\includegraphics[width=0.5\textwidth,trim={1.25cm 0.8cm 0.25cm 1.25cm},clip]{chapter5/figs/comp/fun9/meshs_l2_pdf_error_formatted.pdf}
	\caption{KS error between the true CDF and the approximations (left) and \(L^2\) error between the true PDF, \(cos(4\pi(x+0.5))+1\), and the approximations for the DG method (blue solid line with crosses), uniformisation method (orange dashed line with diamonds) and QBD-RAP method (green dash-dotted line with circles).}
	\label{fig: fun 9 comp} 
\end{figure}

\FloatBarrier

\section{Travelling wave}
Here we investigate the performance the different schemes for approximating transient distributions of one-dimensional travelling wave problems with varying initial conditions. Consider a (trivial) fluid queue with one phase, generator \(\bs T=[0]\) and rate \(c=1\). The PDE (if/when it exists) which describes this system is 
\[\cfrac{\partial}{\partial t} f(x,t) = -\cfrac{\partial}{\partial x} f(x,t),\]
where \(f(x,t)\) is the density at time \(t\). Given an initial condition, \(f(x,0)\), solutions to this problem are given by 
\[f(x,t) = f(x-t,0)\]
so the solution at time \(t\) is just a shift in the initial condition \(t\) units to the right. We suppose that the fluid queue is bounded, with a lower boundary \(x=0\) and upper boundary \(x=10\).

We use the QBD-RAP, uniformisation and DG schemes to discretise the solution in space. We discretise the interval \([0,10]\) into \(10\) cell, each of width 1. We use \(10,001\) points to approximate the integrals which appear in the construction of the initial conditions, to approximate the integrals appearing in the error metrics, and also as a set of discrete points on which evaluate the CDFs to approximate the KS metric. We use the SSPRK4 method to integrate over time using a step size of \(0.005\), and we evolve the system until time \(t=4\). For the DG method we also implement the Generalise MUSCL (REF, appendix?) slope limiter to stabilise the integration.

First consider the initial condition with PDF \(1(x<1)\) so the level of the fluid queue is uniformly distributed over the first cell. For the DG and uniformisation methods, the initial condition can be represented exatly, whereas, for the QBD-RAP method, it cannot. Thus, in this case, there is no discretisation error in constructing the initial conition for the DG and uniformisation methods.
